Triptych

### 2019-10-23 paper/main.tex
Paper line endings
SHVZK proof and corresponding algorithm update

\subsection{Warning}
There is currently no security model established for Triptych, and no proofs of security (except for correctness). This document currently serves only to describe the construction and its algorithms. It should not be considered secure or safe for use without extensive additional formalization and analysis. This is very early research that is still under active development.

@@

\subsection{Notation}
For a three-dimensional tensor $f \equiv (f_{i,j,k}) \subset \F^3$ and blinding factor $r \in \F$, define the Pedersen tensor commitment $$\com(f,r) \equiv rH + \sum_{i,j,k} f_{i,j,k} G_{i,j,k}$$ using fixed independent generators as described above. Operations on tensors are assumed to be performed componentwise. The ordering of elements of $f$ is assumed to be known and canonical.

@@

\subsection{Prover}
The prover algorithm accepts the following:
\begin{itemize}
\item a set of outputs $\{M_k\}_{k=0}^{N-1} \subset \G$
\item a set of corresponding amount commitments $\{P_k\}_{k=0}^{N-1} \subset \G$
\item a set of private keys $\{r^{(u)}\}_{u=0}^{w-1} \subset \F$
\item a set of commitment amounts and blinding factors $\{(s^{(u)},a_u)\}_{u=0}^{w-1} \subset \F^2$
\item a set $\{l^{(u)}\}_{u=0}^{w-1} \subset \F$ such that:
\begin{itemize}
\item each $M_{l^{(u)}} = r^{(u)}G$ has its private key known to the prover
\item each $P_{l^{(u)}} = s^{(u)}G + a_uH$ has its amount and blinding factor known to the prover
\end{itemize}
\item a set of new amount commitments $\{Q_j\}_{j=0}^{T-1} \subset \G$
\item a set of commitment amounts and blinding factors $\{(t_j,b_j)\}_{j=0}^{T-1} \subset \F^2$ such that:
\begin{itemize}
\item each $Q_j = t_jG + b_jH$ has its amount and blinding factor known to the prover
\item $\sum_{u=0}^{w-1} a_j - \sum_{j=0}^{T-1} b_j = 0$
\end{itemize}
\end{itemize}

The prover then performs these steps:
\begin{itemize}
\item Define $\left\{J^{(u)}\right\}_{u=0}^{w-1} \subset \G$ such that $J^{(u)} \equiv (r^{(u)})^{-1}\hp(M_{l_u})$.
\item Select random $r_A \in \F$ and $\left\{a^{(u)}_{j,i}\right\}_{i=1,j,u=0}^{n-1,m-1,w-1} \subset \F^3$. Define $A \equiv \com(a,r_A)$.
\item Define $\left\{\sigma^{(u)}_{j,i}\right\}_{i,j,u=0}^{n-1,j-1,w-1} \subset \F^3$ such that $\sigma^{(u)}_{j,i} \equiv \delta\left(l^{(u)}_j,i\right)$ (using our decomposition notation), and choose random $r_B \in \F$. Define $B \equiv \com(\sigma,r_B)$.
\item Select random $r_C \in \F$, and define $C \equiv \com(a(1-2\sigma), r_C)$.
\item Select random $r_D \in \F$, and define $D \equiv \com(-a^2, r_D)$.
\item Select random $\left\{\rho^{(u)}_j\right\}_{j,u=0}^{m-1,w-1} \subset \F^2$.
\item For each $0 \leq u < w$, define coefficients $\left\{p^{(u)}_{k,j}\right\}_{k,j=0}^{N-1,m-1}$ such that $$p^{(u)}_k(x) \equiv \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} = \delta\left(l^{(u)},k\right)x^m + \sumj p^{(u)}_{k,j}x^j$$ (using our decomposition notation). Then define $p_{k,j} \equiv \sumu p^{(u)}_{k,j}$ and $p_k(x) \equiv \sumu p^{(u)}_k(x)$ accordingly.
\item Define $\{X_j\}_{j=0}^{m-1} \subset \G$ such that: $$X_j \equiv \sumk p_{k,j}M_k + \sumu \rho^{(u)}_jG$$
\item Define $\{Y_j\}_{j=0}^{m-1} \subset \G$ such that: $$Y_j \equiv \sumk p_{k,j}\hp(M_k) + \sumu \rho^{(u)}_jJ^{(u)}$$
\item Define $\{Z_j\}_{j=0}^{m-1} \subset \G$ such that: $$Z_j \equiv \sumk p_{k,j}P_k + \sumu \rho^{(u)}_jG$$
\item Define the Fiat-Shamir transcript challenge $x \in \F$ computed using all public parameters and these elements (with set notation dropped for brevity): $$M,P,Q,J,A,B,C,D,X,Y,Z$$
\item Define $\left\{f^{(u)}_{j,i}\right\}_{i=1,j,u=0}^{n-1,m-1,w-1}$ such that $f^{(u)}_{j,i} \equiv \sigma^{(u)}_{j,i}x + a^{(u)}_{j,i}$.
\item Define $z_A \equiv r_A + xr_B$ and $z_C \equiv xr_C + r_D$.
\item Define $\left\{z^{(u)}_R\right\}_{u=0}^{w-1} \subset \F$ such that: $$z^{(u)}_R \equiv r^{(u)}x^m - \sumj \rho^{(u)}_jx^j$$
\item Define: $$z_S \equiv x^m\left( \sumu s^{(u)} - \sum_{j=0}^{T-1} t_j \right) - \sumj \left(x^j \sumu \rho^{(u)}_j\right)$$
\end{itemize}

@@

\section{Security}
We are interested initially to know whether the Triptych proving system is correct.

\subsection{Correctness}
To show that a correct proof will always be accepted by an honest verifier, we examine each verifier check separately.

The $A/B$ check passes using the identity $$\sumi \sigma^{(u)}_{j,i} = 1$$ for all $0 \leq j < m$ and $0 \leq u < w$.

The $C/D$ check passes similarly, using the identity $$\left(\sigma^{(u)}_{j,i}\right)^2 = \sigma^{(u)}_{j,i}$$ for all $0 \leq j < m$ and $0 \leq u < w$.

@@

Finally, the $Z$-check:
\begin{eqnarray*}
&& \sumk P_k \left[ \sumu \left( \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} \right) \right] - \sumj x^jZ_j - x^m\sum_{j=0}^{T-1} Q_j - z_SG \\
&=& \sumk P_k p_k(x) - \sumj x^j \left( \sumk p_{k,j}P_k + \sumu \rho^{(u)}_jG \right) - x^m\sum_{j=0}^{T-1} Q_j - z_SG \\
&=& \sumk P_k \left( p_k(x) - \sumj x^j p_{k,j} \right) - \sumj x^j \sumu \rho^{(u)}_jG - x^m\sum_{j=0}^{T-1} Q_j - z_SG \\
&=& x^m\sumu (s^{(u)}G + a_uH) - \sumj x^j \sumu \rho^{(u)}_jG - x^m\sum_{j=0}^{T-1} (t_jG + b_jH) - x^m\left( \sumu s^{(u)} - \sum_{j=0}^{T-1} t_j \right)G + \sumj x^j \sumu \rho^{(u)}_jG \\
&=& \left( \sumu a_u - \sum_{j=0}^{T-1} b_j \right)H \\
&=& 0
\end{eqnarray*}

>>>>

\subsection{Warning}
There is currently no security model established for Triptych, and incomplete proofs of security.
This document currently serves only to describe the construction and its algorithms.
It should not be considered secure or safe for use without extensive additional formalization and analysis.
This is very early research that is still under active development.

@@

\subsection{Notation}
For a three-dimensional tensor $f \equiv (f_{i,j,k}) \subset \F$ and blinding factor $r \in \F$, define the Pedersen tensor commitment $$\com(f,r) \equiv rH + \sum_{i,j,k} f_{i,j,k} G_{i,j,k}$$ using fixed independent generators as described above.
Operations on tensors are assumed to be performed componentwise.
The ordering of elements of $f$ is assumed to be known and canonical.

@@

\subsection{Prover}
The prover algorithm accepts the following:
\begin{itemize}
\item a set of outputs $\{M_k\}_{k=0}^{N-1} \subset \G$
\item a set of corresponding amount commitments $\{P_k\}_{k=0}^{N-1} \subset \G$
\item a set of private keys $\{r^{(u)}\}_{u=0}^{w-1} \subset \F$
\item a set of commitment amounts and blinding factors $\{(s^{(u)},a_u)\}_{u=0}^{w-1} \subset \F$
\item a set $\{l^{(u)}\}_{u=0}^{w-1} \subset \F$ such that:
\begin{itemize}
\item each $M_{l^{(u)}} = r^{(u)}G$ has its private key known to the prover
\item each $P_{l^{(u)}} = s^{(u)}G + a_uH$ has its amount and blinding factor known to the prover
\end{itemize}
\item a set of new amount commitments $\{Q_j\}_{j=0}^{T-1} \subset \G$
\item a set of commitment amounts and blinding factors $\{(t_j,b_j)\}_{j=0}^{T-1} \subset \F$ such that:
\begin{itemize}
\item each $Q_j = t_jG + b_jH$ has its amount and blinding factor known to the prover
\item $\sum_{u=0}^{w-1} a_j - \sum_{j=0}^{T-1} b_j = 0$
\end{itemize}
\end{itemize}

The prover then performs these steps:
\begin{itemize}
\item Define $\left\{J^{(u)}\right\}_{u=0}^{w-1} \subset \G$ such that $J^{(u)} \equiv (r^{(u)})^{-1}\hp(M_{l_u})$.
\item Select random $r_A \in \F$ and $\left\{a^{(u)}_{j,i}\right\}_{i=1,j,u=0}^{n-1,m-1,w-1} \subset \F$.
Define $A \equiv \com(a,r_A)$.
\item Define $\left\{\sigma^{(u)}_{j,i}\right\}_{i,j,u=0}^{n-1,j-1,w-1} \subset \F$ such that $\sigma^{(u)}_{j,i} \equiv \delta\left(l^{(u)}_j,i\right)$ (using our decomposition notation), and choose random $r_B \in \F$.
Define $B \equiv \com(\sigma,r_B)$.
\item Select random $r_C \in \F$, and define $C \equiv \com(a(1-2\sigma), r_C)$.
\item Select random $r_D \in \F$, and define $D \equiv \com(-a^2, r_D)$.
\item Select random $\left\{\rho^{(u)}_j\right\}_{j,u=0}^{m-1,w-1}, \left\{\overline{\rho}^{(u)}_j\right\}_{j,u=0}^{m-1,w-1} \subset \F$.
\item For each $0 \leq u < w$, define coefficients $\left\{p^{(u)}_{k,j}\right\}_{k,j=0}^{N-1,m-1}$ such that $$p^{(u)}_k(x) \equiv \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} = \delta\left(l^{(u)},k\right)x^m + \sumj p^{(u)}_{k,j}x^j$$ (using our decomposition notation).
Then define $p_{k,j} \equiv \sumu p^{(u)}_{k,j}$ and $p_k(x) \equiv \sumu p^{(u)}_k(x)$ accordingly.
\item Define $\{X_j\}_{j=0}^{m-1} \subset \G$ such that: $$X_j \equiv \sumk p_{k,j}M_k + \sumu \rho^{(u)}_jG$$
\item Define $\{Y_j\}_{j=0}^{m-1} \subset \G$ such that: $$Y_j \equiv \sumk p_{k,j}\hp(M_k) + \sumu \rho^{(u)}_jJ^{(u)}$$
\item Define $\{Z_j\}_{j=0}^{m-1} \subset \G$ such that: $$Z_j \equiv \sumk p_{k,j}P_k + \sumu \overline{\rho}^{(u)}_jG$$
\item Define the Fiat-Shamir transcript challenge $x \in \F$ computed using all public parameters and these elements (with set notation dropped for brevity): $$M,P,Q,J,A,B,C,D,X,Y,Z$$
\item Define $\left\{f^{(u)}_{j,i}\right\}_{i=1,j,u=0}^{n-1,m-1,w-1}$ such that $f^{(u)}_{j,i} \equiv \sigma^{(u)}_{j,i}x + a^{(u)}_{j,i}$.
\item Define $z_A \equiv r_A + xr_B$ and $z_C \equiv xr_C + r_D$.
\item Define $\left\{z^{(u)}_R\right\}_{u=0}^{w-1} \subset \F$ such that: $$z^{(u)}_R \equiv r^{(u)}x^m - \sumj \rho^{(u)}_jx^j$$
\item Define: $$z_S \equiv x^m\left( \sumu s^{(u)} - \sum_{j=0}^{T-1} t_j \right) - \sumj \left(x^j \sumu \overline{\rho}^{(u)}_j\right)$$
\end{itemize}

@@

\section{Security}
We are interested initially to show that Triptych proving system is perfectly complete, special honest verifier zero-knowledge, and special sound. Note that being special honest verifier zero-knowledge implies that the construction is also witness-indistinguishable \cite{cramer}.

\subsection{Perfect completeness}
To show that a valid proof will always be accepted by an honest verifier, we examine each verifier check separately.

The $A/B$ check succeeds on a valid proof using the identity $$\sumi \sigma^{(u)}_{j,i} = 1$$ for all $0 \leq j < m$ and $0 \leq u < w$.

The $C/D$ check succeeds similarly, using the identity $$\left(\sigma^{(u)}_{j,i}\right)^2 = \sigma^{(u)}_{j,i}$$ for all $0 \leq j < m$ and $0 \leq u < w$.

@@

Finally, the $Z$-check:
\begin{eqnarray*}
&& \sumk P_k \left[ \sumu \left( \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} \right) \right] - \sumj x^jZ_j - x^m\sum_{j=0}^{T-1} Q_j - z_SG \\
&=& \sumk P_k p_k(x) - \sumj x^j \left( \sumk p_{k,j}P_k + \sumu \overline{\rho}^{(u)}_jG \right) - x^m\sum_{j=0}^{T-1} Q_j - z_SG \\
&=& \sumk P_k \left( p_k(x) - \sumj x^j p_{k,j} \right) - \sumj x^j \sumu \overline{\rho}^{(u)}_jG - x^m\sum_{j=0}^{T-1} Q_j - z_SG \\
&=& x^m\sumu (s^{(u)}G + a_uH) - \sumj x^j \sumu \overline{\rho}^{(u)}_jG - x^m\sum_{j=0}^{T-1} (t_jG + b_jH) - x^m\left( \sumu s^{(u)} - \sum_{j=0}^{T-1} t_j \right)G + \sumj x^j \sumu \overline{\rho}^{(u)}_jG \\
&=& \left( \sumu a_u - \sum_{j=0}^{T-1} b_j \right)H \\
&=& 0
\end{eqnarray*}

\subsection{Special honest verifier zero-knowledge}
We construct a special honest verifier zero-knowledge simulator that, given a random verifier challenge $x$, can construct a proof transcript with identical distribution to a valid proof.

First, observe that the simulator presented in the proof of Lemma 1 in \cite{bootle} translates nearly identically to our setting, since we can flatten tensor commitments to a matrix structure with the required sum property.
If the simulator chooses $B \in \G$ randomly, the cited lemma assures us a valid simulation of the proof elements $A,C,D,z_A,z_C,\{f^{(u)}_{j,i \neq 0}\}$; we may compute each $f^{(u)}_{j,0}$ from here.
Further, in a valid proof, $B$ is independent and uniformly distributed as well.

The decisional Diffie-Hellman assumption implies that the elements $\{X_j\}_{j=1}^{m-1}, \{Y_j\}_{j=1}^{m-1}, \{Z_j\}_{j=1}^{m-1}$ are independent and uniformly distributed in a valid proof, so the simulator may choose these randomly.
The verification checks require that $X_0,Y_0,Z_0$ be uniquely determined by the other elements in the corresponding sets in both real proofs and by the simulator.

Finally, the elements $\{z^{(u)}_R\}_{u=0}^{w-1}$ and $z_S$ are independent and uniformly distributed in valid proofs given random $x$, so the simulator may choose them randomly.
Hence the construction is special honest verifier zero-knowledge.
